Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-02-04 appendix.qmd
Tried 5-fold cross validation to compute mse and investigate the distribution (not specific change with cross validation)

>>>>

<!--
```{r}
#| warning: false
#| echo: false

## UMAP with n_neighbor: 7
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names

eval_data_test <- dplyr::bind_rows(vec)[0, ]
eval_data_test <- eval_data_test |>
  dplyr::mutate_if(is.character, as.numeric)

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)

eval_data_test <- eval_data_test |>
  mutate(data_type = "test")

MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 7")
```

## UMAP with n_neighbor: 15
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names

eval_data_test <- dplyr::bind_rows(vec)[0, ]
eval_data_test <- eval_data_test |>
  dplyr::mutate_if(is.character, as.numeric)

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)

eval_data_test <- eval_data_test |>
  mutate(data_type = "test")

MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 15")
```

## UMAP with n_neighbor: 32
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names

eval_data_test <- dplyr::bind_rows(vec)[0, ]
eval_data_test <- eval_data_test |>
  dplyr::mutate_if(is.character, as.numeric)

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)

eval_data_test <- eval_data_test |>
  mutate(data_type = "test")

MSE_df_umap_3 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 32")
```

## UMAP with n_neighbor: 50
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names

eval_data_test <- dplyr::bind_rows(vec)[0, ]
eval_data_test <- eval_data_test |>
  dplyr::mutate_if(is.character, as.numeric)

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)

eval_data_test <- eval_data_test |>
  mutate(data_type = "test")

MSE_df_umap_4 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 50")
```

MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
MSE_df_umap$param <- factor(MSE_df_umap$param, levels = c("n_neighbors: 7", "n_neighbors: 15", "n_neighbors: 32", "n_neighbors: 50"))

## To draw with AIC
aic_plot_param <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                         y = total_error,
                                                                         color = param
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                               y = total_mse,
                               color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
  ylab("MSE") +
  xlab("Total number of bins")

mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
                                                                                             y = total_mse,
                                                                                             color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset. What is the best parameter choice in UMAP to represent the original data in 2D?"
#| label: fig-diagnosticpltDiffParam
#| out-width: 100%

aic_plot_param + mse_plot_param_training + mse_plot_param_test  +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```
-->

<!--model for S-curve with 5 methods-->
<!--
```{r}
#| warning: false
#| echo: false

tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")

## tSNE

num_bins_tsne_s_curve <- 8
shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
                                                          x = tSNE1, y = tSNE2) ## 1.259938
## To extract bin centroids
hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)

df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data

tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)

## Triangulate bin centroids
tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)

# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_tsne_s_curve <- cal_2d_dist(.data = tr_from_to_df_tsne_s_curve)

## To find the benchmark value
benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)

trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)

trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)

num_bins_umap_s_curve <- 6
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

## Compute 2D distances
distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)

# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)

## TriMAP

num_bins_trimap_s_curve <- 6
shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)

df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data

TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)

## Triangulate bin centroids
tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)

# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_trimap_s_curve <- cal_2d_dist(.data = tr_from_to_df_trimap_s_curve)

## To find the benchmark value
benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)

trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)

trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.25) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)

## PacMAP

num_bins_pacmap_s_curve <- 10
shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)

df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data

PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)

## Triangulate bin centroids
tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)

# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_pacmap_s_curve <- cal_2d_dist(.data = tr_from_to_df_pacmap_s_curve)

## To find the benchmark value
benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)

trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)

trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)

## PHATE

num_bins_phate_s_curve <- 8
shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
                                                           x = PHATE1, y = PHATE2) ## 1.259938
## To extract bin centroids
hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)

df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data

## Identify bins with low-density
identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)

df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
  filter(!(hexID %in% identify_rm_bins))

PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)

## Triangulate bin centroids
tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)

# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_phate_s_curve <- cal_2d_dist(.data = tr_from_to_df_phate_s_curve)

## To find the benchmark value
benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)

trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)

trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Is there a good model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
#| out-width: 100%

trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
  #plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 5) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```
-->
